### 2022-11-30 garbage_collect/main.tex
update psudeocode

\Statex $\diamondsuit$ \tt{\sl{Local}}
\begin{itemize}
\item \tt{\sl{Node} leaf:} \sf{ process's leaf in the tree.}
\item {\color{red}\tt{\sl{int} garbageCollectRound}}
\end{itemize}

@@

\Statex $\triangleright$ Returns the response to $D_i(root,b)$, the \nf{i}th \nf{Dequeue} in \nf{root.blocks[b]}.
\Function{element}{FindResponse}{\sl{int} b, \sl{int} i}
\If{\tt{ root.blocks[b-1].\size}\tt{ + root.blocks[b].num\sub{enq} - i $<$ 0}} \label{checkEmpty}\cmt{Check if the queue is empty.}
\State \tt{lastDequeudFrom.update(b)}
\State \Return \tt{null} \label{returnNull}
\Else \cmt{The response is $E_e(root)$, the \nf{e}th \nf{Enqueue} in the root.}
\State \tt{e= i + (root.blocks[b-1].sum\sub{enq}-root.blocks[b-1].size)} \label{computeE}
\State {\color{red}\tt{<x, y>= root.\Call{DoublingSearch}{e, b}}}
\State {\color{red}\tt{lastDequeudFrom.update(x)}}
\State \Return \tt{root.GetEnqueue(x,y)}\label{findAnswer}
\EndIf
\EndFunction{FindResponse}

@@

\Statex $\leadsto$ \textsf{Precondition: \tt{blocks[start..end]} contains a block with \tt{sum\sub{enq}} greater than or equal to \tt{x}}
\Statex $\triangleright$ \textmd{\color{red} Update needed: search on RBT}.
\Function{int}{BinarySearch}{\sl{int} x, \sl{int} start, \sl{int} end}
% \State \Return \tt{min\{j: blocks[j].sum\sub{enq}$\geq$x\}}
\While{\nf{start<end}}
\State \tt{\sl{int}} \tt{mid= floor((start+end)/2)}
\If{\nf{blocks[mid].sum\sub{enq}<x}}
\State \nf{start= mid+1}
\Else
\State \nf{end= mid}
\EndIf
\EndWhile
\State\Return \nf{start}
\EndFunction{BinarySearch}

@@

\Function{<int, int>}{DoublingSearch}{\sl{int} e, \sl{int} end}
\State \tt{start= end-1} \label{dsearchStart}
\While{\tt{root.blocks[start].sum\sub{enq}}$>=$\tt{e}}
\State \tt{start= max(start-(end-start), 0)} \label{doubling}
\EndWhile \label{dsearchEnd}
\State \tt{b= root.BinarySearch(e, start, end)} \label{dsearchBinarySearch}
\State \tt{i= e- root.blocks[b-1].sum\sub{enq}} \label{DSearchComputei}
\State\Return \tt{<b,i>}
\EndFunction{DoublingSearch}
\end{algorithmic}
\end{algorithm}

@@

\Function{void}{Append}{\sl{block} B} \cmt{Only called by the owner of the leaf.}
{\color{red}\If{\tt{root.head\%p\textsuperscript{2}<p and garbageCollectRound<floor(root.head/p\textsuperscript{2})}}
\State \Call{Help()}{}
\State \tt{root.}\Call{FreeMemory()}{lastDequeuedFrom.Get()-1}
\State \tt{garbageCollectRound=floor(root.head/p\textsuperscript{2})}
\EndIf}
\State \tt{blocks[\head]= B} \label{appendLeaf}
\State \tt{\head= \head+1} \label{appendEnd} 
\State \tt{parent.}\Call{Propagate()}{} 
\EndFunction{Append}

@@

\Statex $\triangleright$ \textmd{Creates a block containing new operations of \nf{this.}children, and then tries to append it to \nf{this}.}
\Function{boolean}{Refresh()}{}
\State \tt{h= \head} \label{readHead}
\ForEach{\tt{dir} {\keywordfont{in}} \tt{\{left, right\}}} \label{startHelpChild1}
\State \tt{h\sub{dir}= dir.\head} \label{readChildHead}
\If{\nf{dir.blocks[h\sub{dir}]!=null}} \label{ifHeadnotNull}
\State{\tt{dir.\Call{Advance}{h\sub{dir}}}} \label{helpAdvance}
\EndIf
\EndFor \label{endHelpChild1}
\State \tt{new= \Call{CreateBlock}{h}} \label{invokeCreateBlock}
\If{\tt{new.num==0}} \Return{\tt{true}} \label{addOP} 
\EndIf
\State{\tt{result= blocks[h].CAS(null, new)}} \label{cas}
\State{\tt{this.\Call{Advance}{h}}} \label{advance}
\State \Return{ \tt{result}}

@@

@@

\Statex $\leadsto$ \textsf{Precondition:~\tt{blocks[b].num\sub{enq}$\geq$i$\geq 1$}}
\Function{element}{GetEnqueue}{\sl{int} b, \sl{int} i} \cmt{Returns the \tt{element} of $E_\tt{i}(\tt{this},\tt{b})$.}
\If{\tt{this} \bf{is} \tt{leaf}}
\State\Return \tt{blocks[b].element} \label{getBaseCase}
\ElsIf{\tt{i <= blocks[b].num\sub{enq-left}}} \label{leftOrRight} \cmt{$E_\tt{i}(\tt{this},\tt{b})$ is in the left child of this node.}
\State \tt{subblockIndex= left.BinarySearch(i+blocks[b-1].sum\sub{enq-left}, blocks[b-1].end\sub{left}+1,} \label{leftChildGet}
\Statex \hspace{10.7em}\tt{blocks[b].end\sub{left})} 
\State \Return\tt{left.}\Call{GetEnqueue}{subblockIndex, i} 
\Else
\State \tt{i= i-blocks[b].num\sub{enq-left}}
\State\tt{subblockIndex= right.BinarySearch(i+blocks[b-1].sum\sub{enq-right}, blocks[b-1].end\sub{right}+1,} \label{rightChildGet}
\Statex \hspace{11.1em}\tt{blocks[b].end\sub{right})} 
\State \Return\tt{right.}\Call{GetEnqueue}{subblockIndex, i} 
\EndIf
\EndFunction{GetEnqueue}

\Statex
\Statex $\leadsto$ \textsf{Precondition: \tt{b}th block of the node has propagated up to the root and \tt{blocks[b].num\sub{deq}$\geq$i}.}
\Function{<int, int>}{IndexDequeue}{\sl{int} b, \sl{int} i} \cmt{\color{red} Update needed: return null when superblock in the root was not found.}
\If{\tt{this} \bf{is} \tt{root}}
\State\Return \tt{<b, i>} \label{indexBaseCase}
\Else
\State \tt{dir= (parent.left==n ? left: right)} 
\State \tt{superblockIndex= parent.blocks[blocks[b].super].sum\sub{deq-dir} > blocks[b].sum\sub{deq} ? \label{computeSuper}
\Statex \hspace{11.3em} blocks[b].super: blocks[b].super+1}

@@

\begin{algorithm}
\caption{\tt{\sl{\color{red}MaxByProcess}}}
\begin{algorithmic}[1]
\setcounter{ALG@line}{121}

@@

\begin{algorithm}
\caption{\tt{\sl{Tree}}}
\begin{algorithmic}[1]
\setcounter{ALG@line}{130}

@@

\begin{algorithm}
\caption{\tt{\sl{Node}}}
\begin{algorithmic}[1]
\setcounter{ALG@line}{140}

\Function{}{FreeMemory}{\sl{int} \tt{b}}
\If{not leaf}
\State \tt{left.FreeMemory(blocks[i].end\sub{left})}
\State \tt{right.FreeMemory(blocks[i].end\sub{right})}
\EndIf
\State \tt{\bf{while} !blocks.CAS(blocks.splitGreater(i))}
\EndFunction{FreeMemory}

@@

\begin{algorithm}
\caption{\tt{\sl{PBRT}}}
\begin{algorithmic}[1]
\setcounter{ALG@line}{140}

@@

\Function{}{GetByBlock}{\sl{int} \tt{i}}
\State \Return \tt{rbt.get(i)}
\If{not found}
\State \tt{return written response}
\EndIf
\EndFunction{FreeMemory}

>>>>

\Statex $\diamondsuit$ \tt{\sl{Local}}
\begin{itemize}
\item \tt{\sl{Node} leaf:} \sf{ process's leaf in the tree.}
% \item {\color{red}\tt{\sl{int} garbageCollectRound}}
\end{itemize}

@@

\Statex $\triangleright$ Returns the response to $D_i(root,b)$, the \nf{i}th \nf{Dequeue} in \nf{root.blocks[b]}.
\Function{element}{FindResponse}{\sl{int} b, \sl{int} i}
\If{\tt{ root.blocks[b-1].\size}\tt{ + root.blocks[b].num\sub{enq} - i $<$ 0}} \label{checkEmpty}\cmt{Check if the queue is empty.}
\State \tt{ \color{red} lastDequeudFrom.update(b)}
\State \Return \tt{null} \label{returnNull}
\Else \cmt{The response is $E_e(root)$, the \nf{e}th \nf{Enqueue} in the root.}
\State \tt{e= i + (root.blocks[b-1].sum\sub{enq}-root.blocks[b-1].size)} \label{computeE}
\State {\color{red}\tt{<x, y>= root.\Call{DoublingSearch}{e, b}}}
\State {\color{red}\tt{lastDequeudFrom.update(x)}}
\State \Return \tt{root.GetEnqueue(x,y)}\label{findAnswer}
\EndIf
\EndFunction{FindResponse}

@@

\Statex $\leadsto$ \textsf{Precondition: \tt{blocks[start..end]} contains a block with \tt{sum\sub{enq}} greater than or equal to \tt{x}}
\Statex $\triangleright$ \textmd{\color{red} Update needed: search on RBT does not need start and end, we can search over whole the red-black tree.}.
\Function{int}{BinarySearch}{\sl{int} x}
\State \Return \tt{\color{red} min\{j: blocks[j].sum\sub{enq}$\geq$x\}}
% \While{\nf{start<end}}
% \State \tt{\sl{int}} \tt{mid= floor((start+end)/2)}
% \If{\nf{blocks[mid].sum\sub{enq}<x}}
% \State \nf{start= mid+1}
% \Else
% \State \nf{end= mid}
% \EndIf
% \EndWhile
% \State\Return \nf{start}
\EndFunction{BinarySearch}

@@

\Function{<int, int>}{DoublingSearch}{\sl{int} e, \sl{int} end} \cmt{{\color{red} I think with garbage collection, doubling search is not needed any more and a binary search on \nf{root.blocks} would be good enough.}}
% \State \tt{start= end-1} \label{dsearchStart} 
% \While{\tt{root.blocks[start].sum\sub{enq}}$>=$\tt{e}}
% \State \tt{start= max(start-(end-start), 0)} \label{doubling}
% \EndWhile \label{dsearchEnd}
% \State \tt{b= root.BinarySearch(e, start, end)} \label{dsearchBinarySearch}
% \State \tt{i= e- root.blocks[b-1].sum\sub{enq}} \label{DSearchComputei}
% \State\Return \tt{<b,i>}
\EndFunction{DoublingSearch}
\end{algorithmic}
\end{algorithm}

@@

\Function{void}{Append}{\sl{block} B} \cmt{Only called by the owner of the leaf.}
\State \tt{\color{red} blocks.TryAppend(B, head)} \label{appendLeaf}
\State \tt{\head= \head+1} \label{appendEnd} 
\State \tt{parent.}\Call{Propagate()}{} 
\EndFunction{Append}

@@

\Statex $\triangleright$ \textmd{Creates a block containing new operations of \nf{this.}children, and then tries to append it to \nf{this}.}
\Function{boolean}{Refresh()}{}
\State \tt{h= \head} \label{readHead}
\ForEach{\tt{dir} {\keywordfont{in}} \tt{\{left, right\}}} \label{startHelpChild1}
\State \tt{h\sub{dir}= dir.\head} \label{readChildHead}
\If{\nf{dir.blocks[h\sub{dir}]!=null}} \label{ifHeadnotNull}
\State{\tt{dir.\Call{Advance}{h\sub{dir}}}} \label{helpAdvance}
\EndIf
\EndFor \label{endHelpChild1}
\State \tt{new= \Call{CreateBlock}{h}} \label{invokeCreateBlock}
\If{\tt{new.num==0}} \Return{\tt{true}} \label{addOP} 
\EndIf
\State{\tt{\color{red} result= blocks.TryAppend(new, h)}} \label{cas}
\State{\tt{this.\Call{Advance}{h}}} \label{advance}
\State \Return{ \tt{result}}

@@

\Statex
\color{red}
\Function{int}{GetLastDequeuedFrom}{}{\cmt{Returns the index that is safe to remove the blocks before that in the node.}}
\State \tt{x= lastDequeuedFrom.Get()-1}
\State \tt{n= root}
\State \bf{while} \tt{n!=this} \bf{do}
\State \tt{dir= left (if this is in left subtree of n): otherwise dir=right}
\State \tt{    x=n.blocks[x].end\sub{dir}}
\State\bf{end while}
\EndFunction{GetLastDequeuedFrom}

@@

\Statex $\leadsto$ \textsf{Precondition:~\tt{blocks[b].num\sub{enq}$\geq$i$\geq 1$}}
\Function{element}{GetEnqueue}{\sl{int} b, \sl{int} i} \cmt{Returns the \tt{element} of $E_\tt{i}(\tt{this},\tt{b})$.}
\If{\tt{this} \bf{is} \tt{leaf}}
\State\Return \tt{blocks[b].element} \label{getBaseCase}
\ElsIf{\tt{i <= blocks[b].num\sub{enq-left}}} \label{leftOrRight} \cmt{$E_\tt{i}(\tt{this},\tt{b})$ is in the left child of this node.}
\State \tt{subblockIndex= left.BinarySearch(i+blocks[b-1].sum\sub{enq-left}, blocks[b-1].end\sub{left}+1,}  \label{leftChildGet}
\Statex \hspace{10.7em}\tt{blocks[b].end\sub{left})} \cmt{{\color{red} start and end values are not needed anymore?}}
\State \Return\tt{left.}\Call{GetEnqueue}{subblockIndex, i} 
\Else
\State \tt{i= i-blocks[b].num\sub{enq-left}}
\State\tt{subblockIndex= right.BinarySearch(i+blocks[b-1].sum\sub{enq-right}, blocks[b-1].end\sub{right}+1,} \label{rightChildGet}
\Statex \hspace{11.1em}\tt{blocks[b].end\sub{right})} \cmt{{\color{red} start and end values are not needed anymore?}}
\State \Return\tt{right.}\Call{GetEnqueue}{subblockIndex, i} 
\EndIf
\EndFunction{GetEnqueue}

\Statex
\Statex $\leadsto$ \textsf{Precondition: \tt{b}th block of the node has propagated up to the root and \tt{blocks[b].num\sub{deq}$\geq$i}.}
\Function{<int, int>}{IndexDequeue}{\sl{int} b, \sl{int} i} \cmt{{\color{red} Update needed: return null when superblock in the root was not found.}}
\If{\tt{this} \bf{is} \tt{root}}
\State\Return \tt{<b, i>} \label{indexBaseCase}
\Else
\State \tt{dir= (parent.left==n ? left: right)} 
\State \tt{superblockIndex= parent.blocks[blocks[b].super].sum\sub{deq-dir} > blocks[b].sum\sub{deq} ? \label{computeSuper} 
\Statex \hspace{11.3em} blocks[b].super: blocks[b].super+1} \cmt{{\color{red} Preconditions might be not met.}}

@@

\begin{algorithm}
\color{red}
\caption{\tt{\sl{\color{red}MaxByProcess}}}
\begin{algorithmic}[1]
\setcounter{ALG@line}{121}

@@

\begin{algorithm}
\color{red}
\caption{\tt{\sl{Tree}}}
\begin{algorithmic}[1]
\setcounter{ALG@line}{130}

@@

\begin{algorithm}
\color{red}
\caption{\tt{\sl{Node}}}
\begin{algorithmic}[1]
\setcounter{ALG@line}{140}

\Function{}{FreeMemory}{\sl{int} \tt{b}}
\If{not leaf}
\State \tt{left.FreeMemory(blocks[i].end\sub{left}-1)}
\State \tt{right.FreeMemory(blocks[i].end\sub{right}-1)}
\EndIf
\State \tt{blocks= blocks.splitGreater(i)} \cmt{I think CAS is not needed.}
\EndFunction{FreeMemory}

@@

\begin{algorithm}
\color{red}
\caption{\tt{\sl{PBRT}}}
\begin{algorithmic}[1]
\setcounter{ALG@line}{140}

@@

\Statex
\Function{}{GetByBlock}{\sl{int} \tt{i}}
\State \Return \tt{rbt.get(i)}
\If{not found}
\State \tt{return written response}
\EndIf
\EndFunction{GetByBlock}

\Statex
\Function{}{TryAppend}{\sl{block} \tt{B}, \sl{int} \tt{i}}
\Statex {\color{red} Tries to append \nf{B} with key \nf{i} to the red-black tree.}
{\color{red}\If{\tt{i\%p\textsuperscript{2}=0}}
\State \Call{Help()}{}
\State \tt{root.}\Call{FreeMemory()}{lastDequeuedFrom.Get()-1}
\State \tt{garbageCollectRound=floor(root.head/p\textsuperscript{2})}
\EndIf}
\EndFunction{FreeMemory}
